Paper

### 2019-11-18 Paper_resubmit.Rmd
checking snp data against new models

---
title: "Untitled"
author: "Sarah Killcoyne"
date: "07/11/2019"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4
---

@@

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')

@@

@@

getPalette = colorRampPalette(brewer.pal(5, "Blues"))
p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
  geom_line(linetype='solid', size=0.5) + 
  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
  scale_x_reverse() +
  scale_color_manual(values=getPalette(length(unique(df$is)))) +
  labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
p
ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

@@

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
```

@@

ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

@@

@@

@@

```{r}
roc = pROC::roc(Status ~ Probability, data=val.preds, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(roc) + labs(title='Validation cohort')
ggsave('plots/auc/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=6, height=4, units='in',dpi=300)

@@

```{r}
rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
rp
ggsave(filename='plots/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)

@@

>>>>

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2019-11-13"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4
---

@@

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
patient.info = all.patient.info

@@

i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best', transpose=F) %>% as_tibble %>% mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
  rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))

@@

getPalette = colorRampPalette(brewer.pal(5, "Blues"))
p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
  geom_line(linetype='solid', size=0.5) + 
  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
  scale_x_reverse() +
  scale_color_manual(values=getPalette(length(unique(df$is)))) +
  labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
p

blues = RColorBrewer::brewer.pal(5,'Blues')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC*100))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

@@

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC'), width=6, height=6, units="in", limitsize=F, dpi = 300)
```

@@

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

@@

```{r}
roc$model = "Per Sample"

m = rbind.data.frame(
      get.roc.stat(roc),
      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
m$model = factor(m$model)

cols = brewer.pal(11, 'Paired')
dodge = position_dodge(width=0.9)
mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
p

ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
```

@@

```{r, eval=F}
file = paste(cache.dir, 'model_data.Rdata', sep='/')
load(file, verbose=T)

vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
v.tiles = do.call(bind_rows, purrr::map(list.files(vdir, '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
})) %>% dplyr::rename(Samplename = 'X1')
v.tiles = filter(v.tiles, Samplename %in% val.info$Samplename)
v.tiles = v.tiles[-which(duplicated(v.tiles$Samplename))[2],]

v.arm.tiles = do.call(bind_rows, purrr::map(list.files(vdir, 'arms_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
})) %>% dplyr::rename(Samplename = 'X1')
v.arm.tiles = filter(v.arm.tiles, Samplename %in% val.info$Samplename)
v.arm.tiles = v.arm.tiles[-which(duplicated(v.arm.tiles$Samplename))[2],]

for (i in 1:length(z.mean))
  v.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.tiles[,i+1], z.mean[i], z.sd[i])

for (i in 1:length(z.arms.mean))
  v.arm.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.arm.tiles[,i+1], z.arms.mean[i], z.arms.sd[i])

vt = dplyr::select(v.tiles, -Samplename) %>% as.matrix; rownames(vt) = v.tiles$Samplename
at = dplyr::select(v.arm.tiles, -Samplename) %>% as.matrix; rownames(at) = v.arm.tiles$Samplename

v.df = BarrettsProgressionRisk:::subtractArms(vt, at)

valCNdf = as_tibble(t(v.df), rownames='locs') %>% set_names(c('locs',val.info$Samplename)) %>%
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>%   
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  #dplyr::mutate( seg.type = '5MB' ) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(val.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')'))

#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value))

@@

```{r}
roc = pROC::roc(Status ~ Probability, data=val.preds, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(roc) + labs(title='Validation cohort')
ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=6, height=4, units='in',dpi=300)

@@

```{r}
rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
rp
ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=9, height=7, units='in', dpi=300)

@@

# Clinical Story

## Vignettes

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
sum.patient.data = sum.patient.data %>% mutate(smoking_str = recode(smoking, 'Y'='smoker', 'N'=''))
sum.patient.data$smoking_str = replace_na(sum.patient.data$smoking_str, '')

create.desc<-function(pt,age,sex,smk,len,date) {
  str =  paste0('Patient ', pt, ': ', age, sex)
  if (smk != '' & !is.na(smk)) str = paste0(str, ' ,', smk)
  str = paste0(str,'\n')

  if (len != '' & !is.na(len)) str = paste0(str, len,'cm length ')

  str = paste0(str, 'BE diagnosed ', date)

  return(str)
}
sum.patient.data$desc = sum.patient.data %>% group_by(Patient) %>% rowwise() %>% dplyr::summarise(desc = create.desc(Patient, age.diagnosis, gender, smoking_str, M, initial.endo))

for (pt in c(34,56,20,86,3,23,33,49,51,6,25,68)) {
  length = sum.patient.data %>% ungroup %>% filter(Patient == pt) %>% dplyr::summarise(years = final.endo-initial.endo) %>% pull
  height = as.numeric(max( pg.samp %>% ungroup %>% filter(Patient == pt) %>% dplyr::select(Block) %>% pull ) ) + 1

  p = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) 
  vigLeg = get.legend(p)
  p = p + theme(legend.position = 'none')
  ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p, width=length, height=height,dpi=300)
}
ggsave('plots/vignette/legend.png', vigLeg, width=8, height=1, dpi=300)

### Recommendations

 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
 3. Recheck endoscopy 12-24 months: One or more moderate risk predictions
 4. Regular surveillance 3-5 years: Two or more consecutive low risk predictions

```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}

rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
  rules = as.data.frame(matrix(ncol=0, nrow=0))
  # Consecutive
  for (i in 1:length(brks)) {
    risks = table(Risk[i:(i+1)])
    p53 = table(p53.Status[i:(i+1)])
    rule = 'None'
    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
      rule = 1
    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
      rule = 2
    } else if ( risks['Moderate'] > 0 ) {
      rule = 3
    } else if ( risks['Low'] == 2) {
      rule = 4
    }
  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
  }  
  colnames(rules) = c('t1','t2', 'rule')
  rules$Patient = Patient
  rules
}

pg.samp = pg.samp %>% mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T))

maxP = pg.samp %>% dplyr::group_by(Patient, Status, brks) %>% 
  dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=max(Risk), 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )

rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  r
}))

rules$t1 = factor(rules$t1, levels=levels(pg.samp$brks), ordered=T)
rules$t2 = factor(rules$t2, levels=levels(pg.samp$brks), ordered=T)

penult = filter(rules, t2 != 'Endpoint' & (t1 == rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))

penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(penult %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Penult Endoscopy: % patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)

endpoint = subset(rules,  t2 != 'Endpoint') %>% 
  dplyr::group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

endpoint %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(endpoint %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (P) per endoscopy') %>% kable_styling(full_width=F)

path.rx<-function(Patient, brks, Pathology) {
  rules = as.data.frame(matrix(ncol=0, nrow=0))
  # Consecutive
  for (i in 1:length(brks)) {
    #risks = table(Pathology[i:(i+1)])
    rule = 'None'
    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
      rule = 1
    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
      rule = 2
    } else if ( length(which(grepl('BE|NDBE|ID', Pathology[i:(i+1)]))) > 1  ) {
      rule = 3
    }
  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
  }  
  colnames(rules) = c('t1','t2', 'rule')
  rules$Patient = Patient
  rules
}

rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  r
}))
rules2$t1 = factor(rules2$t1, levels=levels(pg.samp$brks), ordered=T)
rules2$t2 = factor(rules2$t2, levels=levels(pg.samp$brks), ordered=T)

second = subset(rules2,  t2 != 'Endpoint') %>% 
  dplyr::group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(second %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (PATH) per endoscopy') %>% kable_styling(full_width=F)
